\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-25 draft.tex
2 and 3

%For presentation purposes, we provide a sketch of a highly simplified language
%for describing reducers. 
%%
%The language allows us to give examples of reducers and
%to discuss our approach for checking their commutativity. 
%
To present our approach, we introduce a highly simplified language
for describing reducers. 
%
Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, a non-deterministically chosen integer value $*$, or a combination of integer expressions over basic
arithmetic operations. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly for
each input element until the entire input list is consumed. An example of a
reducer is shown in Figure~\ref{fig:reducer_example}. In the paper, to simplify
the presentation, we assume that a reducer does never produce any output in the
loop body $s_2$. In the J-ReCoVer tool, an algorithm to deal with the output
inside the loop is implemented (as briefly mentioned at the end of
Section~\ref{sec:prover}).

@@

The input of J-ReCoVer is a reducer program written in Java, which is the most
popular programming language used in the Hadoop MapReduce
framework. The J-ReCoVer tool has three main components, the
\emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name
suggests, the Preprocessor reads as input a reducer program and performs the
required preprocessing. The goal of Prover is to show that a given reducer is
commutative, and the goal of BugFinder is the opposite. The architecture of
J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input
a reducer program to J-ReCoVer either through our web-interface or use a binary
application installed on his/her own machine.

The \emph{Preprocessor component} first compiles a reducer program to bytecode
and uses the tool Soot to further convert it to the so-called ``Jimple'' format,
which is an intermediate language designed to simplify the analysis of Java
programs. Under the Hadoop MapReduce framework, the permutation of the input is
handled by the scheduler/shuffler component and is affected by issues like
network latency, which are not controllable by programmers. In order to deal
with such issues, we wrote our own dummy Hadoop environment for the reducer as a
part of the Preprocessor component so that the input order of the reducer is now
controlled by J-ReCoVer. Finally, the Preprocessor performs  a program transformation to simplify the analysis.

@@

The \emph{BugFinder component} randomly generates pairs of lists. The lists in a
pair are permutations of each other. A concrete counterexample is reported if
inputting any pair of lists from the generated ones to the reducer generates
different results. Our procedure for generating random pairs is quite naive. We
use five different input list of lengths 5, 7, 9, 11, and 13. For each length,
we generate 300 lists and pick uniformly at random one of its permutations.
Although the approach is simple; in practice, it finds counterexamples in all of
our non-commutative benchmarks in a few seconds.

The \emph{Prover component} reduces the commutativity problem to an SMT
satisfiability problem. From a high-level point of view, we check equivalence
between a reducer program and its variant that has two consecutive inputs
swapped. We show that this can be reduced to a first-order formula and give it
to the SMT solver Z3 for solving. In case that Z3 proves the formula
unsatisfiable, we know that swapping any two consecutive inputs of the reducer
will not change its output. Since all permutations of a list can be obtained by
swapping consecutive elements finitely many times, it follows that the reducer
outputs the same value for all permutations of the same list of inputs. In this
case, J-ReCoVer stops and reports that the reducer is commutative.

>>>>

%For presentation purposes, we provide a sketch of a highly simplified language
%for describing reducers. 
%%
%The language allows us to give examples of reducers and
%to discuss our approach for checking their commutativity. 
%
To present our approach, we introduce a highly simplified language
for describing reducers. 
%
Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, a non-deterministically chosen integer value $*$, or a combination of integer expressions over basic
arithmetic operations. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly for
each input element until the entire input list is consumed. An example of a
reducer is shown in Figure~\ref{fig:reducer_example}. In the paper, to simplify
the presentation, we assume that a reducer does never produce any output in the
loop body $s_2$. J-ReCoVer implements an algorithm to deal with an output
inside the loop (as briefly mentioned at the end of
Section~\ref{sec:prover}).

@@

The input of J-ReCoVer is a reducer program written in Java, which is the most
popular programming language used in the Hadoop MapReduce
framework. The J-ReCoVer tool has three main components,
\emph{Preprocessor}, \emph{Prover}, and \emph{BugFinder}. As the name
suggests, Preprocessor reads as input a reducer program and performs the
required preprocessing. The goal of Prover is to show that a given reducer is
commutative, and the goal of BugFinder is the opposite. The architecture of
J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input
a reducer program to J-ReCoVer either through our web-interface or use a binary
application installed on his/her own machine.

The \emph{Preprocessor component} first compiles a reducer program to bytecode
and uses the tool Soot\footnote{\url{https://github.com/Sable/soot}} to further convert it to the so-called ``Jimple'' format,
which is an intermediate language designed to simplify the analysis of Java
programs. Under the Hadoop MapReduce framework, the permutation of the input is
handled by the scheduler/shuffler component and is affected by issues like
network latency, which are not controllable by programmers. In order to deal
with such issues, we wrote our own dummy Hadoop environment for the reducer as a
part of the Preprocessor component so that the input order of the reducer is now
controlled by J-ReCoVer. Finally, the Preprocessor performs  a program transformation to simplify the analysis.

@@

The \emph{BugFinder component} generates random pairs of lists, 
with the list of each pair being permutations of each other. 
A concrete counterexample is reported if
the reducer outputs different results for the two lists of a generated pair. 
Our procedure for generating random pairs is quite naive. We
use five different input list of lengths 5, 7, 9, 11, and 13. For each length,
we generate 300 lists and pick uniformly at random one of its permutations.
Although the approach is simple; in practice, it finds counterexamples in all of
our non-commutative benchmarks in few seconds.

The \emph{Prover component} reduces the commutativity problem to an SMT
problem. From a high-level point of view, we are checking equivalence
between a reducer program and its variant that has two consecutive inputs
swapped. We show that this equivalence check can be reduced to a first-order formula and give it
to the SMT solver Z3 for solving. In case that Z3 proves the formula
unsatisfiable, we know that swapping any two consecutive inputs of the reducer
will not change its output. Since all permutations of a list can be obtained by
swapping consecutive elements finitely many times, it follows that the reducer
outputs the same value for all permutations of the same list of inputs. In this
case, J-ReCoVer stops and reports that the reducer is commutative.
